Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-16 forecast/paper/paper.tex
Edits thru discussion
- Left parts C and D of results blank, until Dan gets figures
- Discussion still needs to be fleshed out
- Figs 1 and 2 still need to be replaced by their real versions

\title{Can Auxiliary Indicators Improve COVID-19 Forecasts and Hotspot
  Prediction?}

@@

% Please add a significance statement to explain the relevance of your work
% TODO write a real one. This is the indicator's one.
\significancestatement{To study the COVID-19 pandemic, its effects on society,
  and measures for reducing its spread, researchers need detailed data on the
  course of the pandemic. Standard public health data streams suffer
  inconsistent reporting and frequent, unexpected revisions. They also miss
  other aspects of a population's behavior, worthy of consideration. We
  present an open database of COVID signals in the United States, measured at
  the county level and updated daily. These include traditionally reported COVID
  cases and deaths, and many others: signals based on measures of mobility,
  social distancing, internet search trends, self-reported symptoms, and
  patterns COVID-related activity in de-identified medical insurance claims. The
  database provides all signals in a common, easy-to-use format, empowering both
  public health research and operational decision-making.
}

@@

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Tracking and Impact Survey CLI-in-community (CTIS-CLI-in-community): 
  The estimated percentage of the population who knows someone who is sick,
  based on Delphi's surveys of Facebook users.
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
  COVID-19 related Google search volume for queries related to anosmia (loss of
  smell), ageusia (loss of taste), or both.  
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

@@

\begin{figure}[tb!]
\centering
\includegraphics[width=\columnwidth]{fig/vintage.pdf}
\caption{TODO}
\label{fig:vintage}
\end{figure}

Vintage training data means different things, in practice, for different
signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
DV-CLI, are most often 3-5 days latent, and subject to a considerable but
regular degree of revision or ``backfill'' after their initial publication date.
The survey-based signal, CTIS-CLI-in-community, is 2 days latent, and rarely
undergoes any revision at all.  The target variable itself, reported COVID-19
case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
(sometimes clear anomalies) after initial publication.  Compared to the revision
pattern in the medical claims signals, which are much more systematic in nature,
revisions in case reports can be highly erratic. Big spikes or other anomalies
can occur in the data as reporting backlogs are cleared, changes in case
definitions are made, etc. Groups like JHU CSSE then work tirelessly to correct
such anomalies after first publication (e.g., they will attempt to
back-distribute a spike when a reporting backlog is cleared, by working with a
local authority to figure out how this should best be done), which can result
in very nontrivial and often unpredictable revisions.  See Figure
\ref{fig:vintage} for an example.

@@

\begin{table*}[t]
  \centering
  \caption{Summary of forecasting and hotspot prediction tasks considered in
    this paper.}
  \begin{tabular}{l p{2.75in} p{2.75in}}
    \toprule
    & \textbf{Forecasting} & \textbf{Hotspot prediction} \\
    \midrule
    Response variable & $Y_{\ell,t}$ (7-day trailing average of COVID-19 case
 incidence rates, per location $\ell$ and time $t$) & $Z_{\ell,t} =
\indicator{Y_{\ell,t} \geq 1.25 \cdot Y_{\ell, t-7}}$ (indicator that $Y_{\ell,t}$ 
grows by more than 25\% relative to the preceding week) \\
    Geographic resolution & Hospital referral region (HRR) & Hospital referral
region (HRR) \\ 
    Forecast period & June 6--December 31, 2020 & June 13--December 31, 2020 \\  
    Model type & Quantile regression & Logistic regression \\
    Evaluation metric & Weighted interval score (WIS) & Area under curve (AUC) \\
    \bottomrule
  \end{tabular}
\label{tab:analysis_tasks}
\end{table*}

@@

In our analysis, we let the prediction date $t$ run over each day in between
early/mid June and December 31, 2020.  The precise start date differs for 
forecasting and hotspots prediction; for each task it was chosen to be the
earliest date at which the data needed to train all models was available, which
ends up being (per our setup, with 21 days of training data, and lagged values
of signals for features as we will detail shortly) June 6, 2020 for forecasting,
and June 13, 2020 for hotspot prediction. (The bottleneck here is the
CTIS-CLI-in-community signal, which does not exist before early April 2020, when
the survey was first launched). The end date was chosen again with a
consideration to align both tasks as best as possible, and because few hotspots
exist post December 31, 2020, due to the general and gradual decline of the
pandemic in 2021.

@@

Recall $Y_{t,\ell}$ denotes the 7-day trailing average of COVID-19 case
incidence rates in location $\ell$ and at time $t$.  Separately for each
$a=7,\ldots,28$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
simple probabilistic forecasting model of the form:     
\begin{equation}
\label{eq:forecast_ar}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s}, s \leq t)  
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
\end{equation}
This model uses current case rates, and the case rates 7 and 14 days ago, in
order to predict (the quantiles of) case rates in the future.  We consider a
total of 7 quantile levels (chosen in accordance with the county-level quantile
levels suggested by the COVID-19 Forecast Hub),      
\begin{equation}
\label{eq:quantile_levels}
\tau \in \{0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975 \}.
\end{equation}
We fit \eqref{eq:forecast_ar} using \textit{quantile regression}
\cite{Koenker:1978, Koenker:2005, Koenker:2006} separately for each $\tau$, 
using data from all 306 HRRs, and within each HRR, using the most recent 21 days
of training data.  This gives us 6,426 training samples for each quantile 
regression problem.

@@

At prediction time, in order to deal with possible crossing violations (a
predicted quantile at level $\tau$ exceeds a predicted quantile at level $\tau'
> \tau$), we apply a simple post-hoc sorting.  See Figure \ref{fig:trajectory}
for an example forecast.

\begin{figure}[tb!]
\centering
\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
\caption{TODO}
\label{fig:trajectory}
\end{figure}

@@

Define the binary indicator:
$$
Z_{t,\ell} = \indicator{Y^\Delta_{\ell,t} \geq 0.25},
$$
where we use the notation \smash{$Y^\Delta_{\ell,t} = (Y_{\ell,t} - Y_{\ell,
    t-7})/(Y_{\ell,t-7})$}. In other words, $Z_{t,\ell}=1$ if the number of
newly reported cases over the past 7 days has increased by at least 25\%
compared to the preceding week.  When this occurs, we say location $\ell$ is a
\textit{hotspot} at time $t$.  Empirically, this rule labels about 27\% of
location-time pairs as hotspots, during the prediction period (June 13--December
31, 2020).

@@

Some signals display a significant degree of spatial heterogeneity, by which we
mean the values that they take at different geographic locations are not 
comparable.  This is true of the Google-AA signal (due to the way in which the 
underlying search trends time series is self-normalized, see
\cite{GoogleSymptoms}) and of the claims-based signals (see the discussion in
our companion paper on the API and database).  Such spatial heterogeneity likely
hurts the performance of the predictive models that rely on these signals,
as we train the models on data pooled over all locations.  In the current 
paper, we do not attempt to address this issue (it is again a topic of ongoing
work in our group), and we simply note that location-specific effects
(or pre-processing to remove spatial bias) would likely improve the performance 
of the models involving Google-AA and the claims-based indicators.

@@

Finally, to quantify the importance of training and making predictions using
proper vintage data, we ran a parallel set of forecasting and hotspot prediction 
experiments using finalized data. The results, given in the supplement, show
that training and making predictions on finalized data can result in grossly
optimistic estimates of true test-time performance (up to 10\% better in terms
of average WIS or AUC).

@@

@@

---which is that 
  defining features from lagged indicator values in the  models in
  \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x} simply provides a kind of
  implicit regularization on top of the basic AR model (so the benefits have
  nothing to do with the actual indicator values)---

To test the ``implicit regularization'' hypothesis (which we explain
  precisely later), we reran all of the prediction experiments but with
  $X_{\ell,t}$ in each indicator model replaced by suitable random noise.  This 
  resulted in vastly different (worse) performance compared to the indicator
  models fit with actual indicator values, and gave no improvement over the 
  pure AR model, in both forecasting and hotspot prediction, providing a strong
  rejection (informally speaking) of the implicit regularization hypothesis.

@@

\subsection{Leading-ness or Lagging-ness}

@@

% TODO: make th

A Primary 
B Bootstrap
C up/down/flat (Dan's first histogram) --> table in appendix
D leading/lagging

\subsection{Primary Results}

Figure \ref{fig:topline-overall-results} displays the primary results
of this paper, showing the effect of using indicators on the two
forecasting tasks as a function of the horizon of the forecast (i.e.,
the number of days ahead one is forecasting). For the case rate forecasting
task, the left panel shows the ratio of the mean
WIS 
of each forecaster relative to the mean WIS of the
baseline. \attn{Further description and justification of this metric
  should be added to the ``Evaluation Metrics'' section rather than
  here it seems?}  Both means are taken over all HRRs and forecast
dates ranging from June 9, 2020 to December 31, 2020 \attn{verify}.
All curves are well below 1, which means that all methods, including \ar, outperform the baseline across all
considered forecast horizons. The claims-based
indicators \dv~and \chngcov~show the most substantial gains, whereas
the \chngcli~does not show any improvement over \ar.  The \fb~and
\gs~indicators also show a non-negligible improvement.  Notably,
\fb~is the only indicator to offer a clear improvement at far aheads.
%\begin{itemize}
%\item \dv~and \chngcov~perform very well
%\item \fb~and \gs~perform well
%\item \chngcli~performs poorly
%\item Interestingly, fb is the only one offering clear improvement at
 % far aheads
%\item Note the y-axis: consistent, clear improvement over baseline
%\end{itemize}

For the hotspot prediction task, the right panel of Figure
\ref{fig:topline-overall-results} shows the AUC of each method.
The monotonic decrease of all curves reflects the fact that hotspot
prediction gets harder as the forecasting horizon increases. At the
highest considered ahead (nearly one month ahead), the AUCs are close
  to 0.5, which is no better than random guessing.
In the 7-14 day ahead range, the indicators show a noticeable
  improvement over \ar.  For larger horizons there does not
  appear to be much difference in methods.
The approximate ordering of methods is similar to what is seen
  in the forecasting task: \chngcov~and \dv~are the strongest
  performing and \ar~the weakest.
Even for 7-day ahead predictions, the AUC is below 0.7,
  suggesting that all methods struggle with this classification
  task. \attn{Do people agree that an AUC of 0.7 is not all that great?}
  That said, \dv~and \chngcov~attain roughly a 10\% increase in AUC
  over the \ar~classifier.

%\begin{itemize}
%\item Monotonic decrease reflects the fact that hotspot
%  prediction gets harder as ahead increases.
%\item At the highest ahead (almost one month ahead) the AUCs are close
%  to 0.5, which is no better than random guessing.
%\item In the 7-14 day ahead range, the indicators show a noticeable
%  improvement over the \ar.  For larger horizons there does not
%  appear to be much difference in methods
%\item The approximate ordering of methods is similar to what is seen
%  in the forecasting task: chng\_cov and dv are the strongest
%  performing and \ar the weakest.
%\item Even for 7-day ahead predictions, the AUC is below 0.7,
%  suggesting that all methods struggle with this classification task.
%  That said, dv and chng\_cov attain roughly a 10\% increase in AUC
%  over the \ar classifier.
%\end{itemize}

% We see that all three \attn{will need to
% include imputed-zero google here, hopefully four} indicator-assisted
% forecasters improve upon the baseline QAR3 forecaster.  The QAR3 model
% assisted by the Change Healthcare confirmed COVID-19 signal
% (QAR3+CHCOV3) sees the strongest improvement.
% C`

\begin{figure}[tb!]
  \centering
  \includegraphics[width=\linewidth]{fig/figure5.pdf}
  \caption{\label{fig:topline-overall-results}  Forecasting performance as a
    function of number of days 
    ahead, aggregated across all HRRs and forecast dates. \attn{How's this?}}
\end{figure}

\subsection{Avoiding misleading retrospective evaluations}

Figure \ref{fig:finalized-vs-honest} quantifies the effect of
not properly accounting for the question of ``what was known when'' in
performing retrospective evaluations of forecasters.  When methods are
given the finalized version of the data rather than the version available at the time that the forecast would have been made, all methods
appear (falsely) to have better performance.  For example, for forecasting case rates
7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
recorded using the finalized values of the data.  This effect
diminishes as the forecasting horizon increases, reflecting the fact
that these forecasters rely less heavily on recent data than very
short-term forecasters.  Crucially, some methods are
``helped'' more than others by the less scrupulous retrospective
evaluation, underscoring the difficulty of avoiding misleading
conclusions when performing retrospective evaluations of forecasters.

The \chngcli~indicator (along with the other claims-based signals) is the most
affected by this distinction, reflecting the latency in claims-based
reporting. \attn{Actually, \dv~is the least affected for forecasting
  whereas it is one of the most affected for hotspots.}  This supports
the importance of efforts to provide ``nowcasts'' for claims signals
(which corresponds to a 0-ahead ``forecast'' of what the claims
signal's value will be once all data has been collected).

Even the \ar~model is affected by this distinction, reflecting the
fact that the case rates themselves (i.e., the response values) are also
subject to revision.  The forecasters based on indicators are thus
affected both by revisions to the indicators and by revisions to the
case rates.

\begin{itemize}
%\item Motivates the importance of nowcasting for claims signals.
  \item For forecasting task: Also, \chngcli~and \dv~perform very similarly here, which is
    reassuring because they're measuring the same thing (in
    principle).  Yet they must have very different backfill profiles…
  \item For hotspots task: dv seems to get a bigger boost than
    \chngcov - interesting dichotomy to forecasting
\item But in both forecasting/hotspots \chngcli~is affected a lot by backfill
\item \attn{We should be careful about which indicators we were not
      able to track latency for.  In fact, we should probably remove
      these from the plot since they would be more strongly affected
      than is shown in the plot.}

\item \attn{We can also discuss the distinction
      between indicators that could in principle have low latency
      (e.g. GS) versus those where latency is inherent
      (e.g. claims-based indicators). If the goal is assessing the
      performance of GS ``for the next pandemic'' then it would make
      sense to give GS a pass since in theory they could have had it
      with very low latency (as opposed to claims based signals where
      backfill may be unavoidable).  Of course, in
    ``the next pandemic'' there might not be such highly specific
    symptoms such as A+A as there was in the COVID-19 case.} 
\end{itemize}

\begin{figure}[tb!]
     \centering
     \includegraphics[width=\linewidth]{fig/figure6.pdf}
     \caption{How misleading is a retrospective analysis that uses the
       finalized version of the data rather than the version that was
       in fact available at the time the forecast was made?  Plots
       show the ratio in performance (using mean WIS fore the
       forecasting task and AUC for the hotspot prediction
       task). \label{fig:finalized-vs-honest} Cheating vs. honest plots.} 
\end{figure}

\subsection{Additional sensitivity analyses}

In addition to whether one properly accounts for signal latency, there
are several other choices in evaluating forecasters that can also have
a non-trivial impact on results.

For example, in Figure
\ref{fig:topline-overall-results}, we show the ratio
of means.  Other choices of aggregation and scaling are also
possible.  For example, taking the mean of ratios of errors is
possible and would give more importance to locations where
errors are small (either because the location-time pairs had low case
rates or because the forecasting task was itself easier).  By
contrast, the ratio of the mean of errors may be dominated by
location-time pairs where the error is large.
Figure \ref{fig:
  scale-aggregate-order} of the Supplementary Information shows...
\begin{itemize}
\item Note the y-axis: barely an improvement over baseline, and this goes away for AR after 11 days ahead, and for all other models around 13-14 days ahead.
\item Using trimmed means here because otherwise the results are extremely unstable (as the denominators can be small, when WIS of baseline is small).
\item Similar message to before (non-adjusted), but \dv~and \chngcov~the best, \fb~and \gs~still good, but now \chngcli~seems to do much better!  
\item What's our interpretation here?  That somehow \chngcli~just
  does particularly bad in tasks where the baseline is also bad?
  \item We might include a histogram of WIS with a log-transformed
    x-axis to suggest that the geometric mean may be reasonable here.
\end{itemize}
Using instead the geometric mean makes the order of aggregation and
scaling immaterial since the ratio of geometric means is the same as
the geometric mean of ratios.  Figure \ref{fig:geometric-mean} of the
Supplementary Information shows how Figure
\ref{fig:topline-overall-results} panel A would change if
one replaces the arithmetic mean by the geometric mean.

In Supplementary Information \ref{sec:2021}, we investigate the effect of lengthening
the time window to include 2021.
\begin{itemize}
\item For forecasting: results get better, whether we look at
  non-adjusted (first plot) or adjusted scores (second plot).  Other
  than that, interpretation is qualitatively similar, except fb now
  seems to be worse than gs.  And \chngcli~is strong, even in the
  non-adjusted view.
  \item For hotspots: Alden will write something explaining why we
    don't do hotspot detection in 2021.
\end{itemize}

In Supplementary Information Section \ref{sec:gs-locations}, we look at various approaches to
accounting for the fact that the \gs~indicator is only available
for XX HRRs.

\begin{itemize}
  \item For forecasting task: clearly gs is not missing at random, and when it's present, it tends to be predictive.  Hence high values have a low false positivity rate.  Most visible in the adjusted view below, where gs triumphs.  Also \chngcli~gets a lot worse.

\item For hotspots: All methods look better, suggesting that the “GS locations” are easier for hotspot prediction. E.g. at 7-days ahead, the AUCs computed based on all locations range from 0.61-0.66; restricted to GS locations, the AUCs range from about 0.65-0.69. (These are all eyeballed, should get actual numbers if we want to say something like this in paper.)
\item gs\_subset appears to be particularly helped by this subsetting
  (which makes sense since on those other locations it was 0-imputed)
  \item Daniel will also try gs\_impute as in forecasting
\end{itemize}

% Figure \ref{fig:forecast-median-relative-wis-faceted} splits the topline
% median relative WIS results over four time periods:  June and July, which
% corresponds to the small, initial surge experienced in the US;
% August and September, when incident cases nationally were somewhat stable;
% October, when a second surge was witnessed; and November and December,
% which saw the largest surges to date.  In essentially all the time periods,
% the indicator-assisted models perform in aggregate as well or better than
% the baseline QAR3 model.  The results in October do not show much
% differentiation between the baseline QAR3 and the indicator-assisted models.
% However, we note that during that time period, even the baseline QAR3
% model is doing very well in terms of relative WIS (especially compared to
% other time periods), so perhaps there is not much ``room for improvement'',
% so to speak.

% \begin{figure*}[tb!]
%      \centering
%      \begin{subfigure}[b]{0.49\textwidth}
%          \centering
%          \includegraphics[width=\textwidth]{fig/reduction_mean_scaled_wis_facet_honest.pdf}
%          \caption{Forecasting task: mean WIS relative to mean WIS of baseline}
%          \label{fig:forecast-reduction-mean-scaled-wis-facet}
%      \end{subfigure}
%      \hfill
%      \begin{subfigure}[b]{0.49\textwidth}
%          \centering
%          \includegraphics[width=\textwidth]{fig/auc_by_forecast_period.pdf}
%          \caption{Hotspot prediction task: AUC}
%          \label{fig:hotspot-auc-faceted}
%      \end{subfigure}
%      \label{fig:topline-faceted-results}
%      \caption{Topline results from June - December 2020, split into
%      four representative time periods.}
% \end{figure*}

\paragraph{\attn{Additional results section material, some for Discussion}}

\begin{figure}
    \centering
    \includegraphics[width=.95\linewidth]{fig/fake-map.pdf}
    \caption{Weighted interval score for forecasts made on April 1, 2021. Scores
      are averaged over horizons from 1 to 14 days ahead and are relative to
      population size.} 
    \label{fig:fake-map}
\end{figure}

\begin{enumerate}
%\item How does this vary over time? Table showing 7-day ahead and 14-day ahead
%  (just picking out two of the d=5,...,20). Include a national curve.
\item Spatial map of errors across HRRs for both tasks. See
  \autoref{fig:fake-map} for an example. 
\item Unlikely we can find it, but we should try anyway. Possibly put in the
  intro or discussion instead. Find an example of a forecast altered by the
  presence of an indicator.  Plot the two trajectories.
\item Examine a specific HRR over time. What does it mean to predict certain
  hotspots? How can we understand what is happening?
 \item What can we say about statistical significance? NRI, paired t-test?
 \item Figures showing hotspots make sense
 \item Horizontal line at 0.5 on AUC curve, discribe that this is random
   guessing. 
\end{enumerate}

@@

We have assessed the performance of [NUMBER] indicators in terms of their
usefulness for performing two forecasting tasks: probabilistic
forecasting of case
rates and hotspot prediction at the HRR-level, 1--[NUMBER] days in advance.
In measuring the practical usefulness of an indicator to forecasting,
we distinguish between honest versus aspirational assessments.  Data
latency is an important factor when considering the usefulness of an
indicator.  Retrospective evaluations that ignore these messy data
issues fail to provide the full picture.  While one can hope that reporting
systems will operate faster and more consistently in ``the next
pandemic,'' some reasons for this latency appear unavoidable.  The
chaos and strain inherent to pandemic times make indicators that can
be reported and collected with quickly with minimal effort or manual
involvement desirable.

We conclude by observing that the
approach we have taken in this paper is somewhat opposite to that of much of the COVID-19
forecasting literature.  We have chosen to consider only very simple forecasting models while devoting most of our effort to accounting for as much of the complexity of
the data and evaluation as possible.  By contrast, many papers focus
on very complicated forecasting approaches but then evaluate them
under unrealistic, retrospective conditions.

\begin{enumerate}

\item Discussion should contain more detailed lit review ("context"), rather
  than in the intro, per the instructions.

\item Use of syndromic surveillance for other diseases?

>>>>

\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
  Prediction?}

@@

% Please add a significance statement to explain the relevance of your work
\significancestatement{
  TODO % For reference, below is the indicators paper one.

  % To study the COVID-19 pandemic, its effects on society,
  % and measures for reducing its spread, researchers need detailed data on the
  % course of the pandemic. Standard public health data streams suffer
  % inconsistent reporting and frequent, unexpected revisions. They also miss
  % other aspects of a population's behavior, worthy of consideration. We
  % present an open database of COVID signals in the United States, measured at
  % the county level and updated daily. These include traditionally reported
  % COVID cases and deaths, and many others: signals based on measures of
  % mobility, social distancing, internet search trends, self-reported symptoms,
  % and patterns COVID-related activity in de-identified medical insurance 
  % claims. The database provides all signals in a common, easy-to-use format,
  % empowering both  public health research and operational decision-making. 
}

@@

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey CLI in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community that is sick, based on Delphi's surveys of
  Facebook users, called the COVID Trends and Impact Survey. 
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
  COVID-19 related Google search volume for queries relating to anosmia or
  ageusia (loss of smell or taste), based on Google's Search Trends data set. 
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

@@

\begin{figure}[tb!]
\includegraphics[width=\columnwidth]{fig/vintage.pdf}
\caption{TODO}
\label{fig:vintage}
\end{figure}

Vintage training data means different things, in practice, for different
signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
DV-CLI, are most often 3-5 days latent, and subject to a considerable but
regular degree of revision or ``backfill'' after their initial publication date.
The survey-based signal, CTIS-CLI-in-community, is 2 days latent, and rarely
undergoes any revision at all.  The target variable itself, reported COVID-19
case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
(sometimes clear anomalies) after initial publication.  Compared to the revision
pattern in the medical claims signals, which are much more systematic in nature,
revisions in case reports can be highly erratic. Big spikes or other anomalies
can occur in the data as reporting backlogs are cleared, changes in case
definitions are made, etc. Groups like JHU CSSE then work tirelessly to correct
such anomalies after first publication (e.g., they will attempt to
back-distribute a spike when a reporting backlog is cleared, by working with a
local authority to figure out how this should best be done), which can result
in very nontrivial and often unpredictable revisions.  See
Figure~\ref{fig:vintage} for an example.

@@

\begin{table*}[t]
\centering
\caption{Summary of forecasting and hotspot prediction tasks considered in
  this paper.}
\begin{tabular}{l p{2.75in} p{2.75in}}
  \toprule
  & \textbf{Forecasting} & \textbf{Hotspot prediction} \\
  \midrule
  Response variable & $Y_{\ell,t}$ (7-day trailing average of COVID-19 case 
incidence rates, per location $\ell$ and time $t$) & $Z_{\ell,t} =
\indicator{Y_{\ell,t} \geq 1.25 \cdot Y_{\ell,t-7}}$ (indicator that 
$Y_{\ell,t}$ grows by more than 25\% relative to the preceding week) \\ 
  Geographic resolution & Hospital referral region (HRR) & Hospital referral
region (HRR) \\ 
  Forecast period & June 9--December 31, 2020 & June 16--December 31, 2020 \\  
  Model type & Quantile regression & Logistic regression \\
  Evaluation metric & Weighted interval score (WIS) & Area under curve (AUC) \\
  \bottomrule
\end{tabular}
\label{tab:analysis_tasks}
\end{table*}

@@

In our analysis, we let the prediction date $t$ run over each day in between
early/mid June and December 31, 2020.  The precise start date differs for 
forecasting and hotspots prediction; for each task it was chosen to be the
earliest date at which the data needed to train all models was available, which
ends up being (per our setup, with 21 days of training data, and lagged values
of signals for features as we will detail shortly) June 9, 2020 for forecasting,
and June 16, 2020 for hotspot prediction. (The bottleneck here is the
CTIS-CLI-in-community signal, which does not exist before early April 2020, when
the survey was first launched). The end date was chosen again with a
consideration to align both tasks as best as possible, and because few hotspots
exist post December 31, 2020, due to the general and gradual decline of the
pandemic in 2021.

@@

Recall $Y_{\ell,t}$ denotes the 7-day trailing average of COVID-19 case
incidence rates in location $\ell$ and at time $t$.  Separately for each
$a=7,\ldots,28$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
simple probabilistic forecasting model of the form:     
\begin{equation}
\label{eq:forecast_ar}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s}, s \leq t)  
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
\end{equation}
This model uses current case rates, and the case rates 7 and 14 days ago, in
order to predict (the quantiles of) case rates in the future.  We consider a
total of 7 quantile levels (chosen in accordance with the county-level quantile
levels suggested by the COVID-19 Forecast Hub),      
\begin{equation}
\label{eq:quantile_levels}
\tau \in \{0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975 \}.
\end{equation}
We fit \eqref{eq:forecast_ar} using \textit{quantile regression}
\cite{Koenker:1978, Koenker:2005, Koenker:2006} separately for each $\tau$, 
using data from all 306 HRRs, and within each HRR, using the most recent 21 days
of training data.  This gives us 6,426 training samples for each quantile 
regression problem.

@@

At prediction time, in order to deal with possible crossing violations (a
predicted quantile at level $\tau$ exceeds a predicted quantile at level $\tau'
> \tau$), we apply a simple post-hoc sorting.  See Figure~\ref{fig:trajectory}
for an example forecast.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
\caption{TODO}
\label{fig:trajectory}
\end{figure}

@@

Define the binary indicator:
$$
Z_{\ell,t} = \indicator{Y^\Delta_{\ell,t} \geq 0.25},
$$
where we use the notation \smash{$Y^\Delta_{\ell,t} = (Y_{\ell,t} - Y_{\ell,
    t-7})/(Y_{\ell,t-7})$}. In other words, $Z_{\ell,t}=1$ if the number of
newly reported cases over the past 7 days has increased by at least 25\%
compared to the preceding week.  When this occurs, we say location $\ell$ is a
\textit{hotspot} at time $t$.  Empirically, this rule labels about 27\% of
location-time pairs as hotspots, during the prediction period (June 16--December
31, 2020).

@@

Some signals have a significant amount of spatial heterogeneity, by which we
mean their values across different geographic locations are not comparable.
This is the case for the Google-AA signal (due to the way in which the
underlying search trends time series is self-normalized, see
\cite{GoogleSymptoms}) and the claims-based signals (see the discussion in our
companion paper on the API and database).  Such spatial heterogeneity likely
hurts the performance of the predictive models that rely on these signals,
because we train the models on data pooled over all locations.  In the current
paper, we do not attempt to address this issue (it is again a topic of ongoing
work in our group), and we simply note that location-specific effects (or
pre-processing to remove spatial bias) would likely improve the performance of
the models involving Google-AA and the claims-based indicators.

@@

Finally, to quantify the importance of training and making predictions using
proper vintage data, we ran a parallel set of forecasting and hotspot prediction 
experiments using finalized data. The results, given in the supplement, show
that training and making predictions on finalized data can result in overly
optimistic estimates of true test-time performance (up to 10\% better in terms
of average WIS or AUC).
% TODO note what we said here about the supplement

@@

Figure~\ref{fig:forecast} displays the results for forecasting, stratified by
ahead value and averaged over all HRRs and forecast dates (June 9--December 31,
2020).  Shown is the average WIS for each forecast model divided by the averaged
WIS from a ``baseline'' model, which is basically a flat-line forecaster (its
median forecast for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted
quantiles defined around this flat line based on historical variation). This is
the same baseline model as that used in the COVID-19 Forecast Hub.  Here, we  
divide by the average WIS of the baseline model in order to put the y-axis on an 
interpretable scale.  We can see that all curves are well below 1, which means
that all of the models, including AR, outperform the baseline on average over
the forecast period (though their advantage over the baseline clearly narrows as
the ahead value grows).

We can also see from the figure that CHNG-COVID and DV-CLI offer the biggest
gains over AR at small ahead values, followed by CTIS-CLI-in-community and
Google-AA, with the former providing the biggest gains at large ahead values.
The CHNG-CLI model performs basically the same as AR.  This is likely due to the
fact that CHNG-CLI suffers from some sort of volatility in forecasting due to
backfill.  The evidence for this is twofold: the CHNG-CLI model benefits from a
more robust method of aggregating WIS (geometric mean; shown in the supplement),
and when we train and make predictions on all-finalized data, it provides a big
gain over AR, on par with the best-performing models (also shown in the
supplement).
% TODO note what we said here about the supplement

\begin{figure}[t]
\includegraphics[width=\columnwidth]{fig/fcast-1.pdf}
\caption{TODO}
\label{fig:forecast}
\end{figure}

Figure~\ref{fig:hotspot} displays the results for hotspot prediction, again
stratified by ahead value and averaged over all HRRs and prediction dates (June 
16--December 31, 2020).  We can see many similarities to the forecasting
results: CHNG-COVID and DV-CLI offer the biggest improvement over AR, all models
including AR degrade in performance towards the baseline (this is now just a
classifier based on random guessing, which achieves an AUC of 0.5) as the ahead
values grow.  However, one clear difference is that the CHNG-CLI model performs
quite well in hotspot prediction, close to the best-performing indicator models
for many of the ahead values. [TODO: should we insert one line explaining
this---you're somewhat protected form volatiliy in a logistic model thanks to
the sigmoid function used to define the probabilities?]

@@

One might ask if the benefits we observe in forecasting and hotspot have
anything to do with the actual indicator themselves; could it instead be that
the indicators are simply providing ``implicit regularization'' on top of the
basic AR model, when we include them as lagged features in
\eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x}?

To test this implicit regularization hypothesis, we reran all of the prediction
experiments but with $X_{\ell,t}$ in each indicator model \eqref{eq:forecast_ar_x},
\eqref{eq:hotspot_ar_x} replaced with suitable random noise (bootstrap
samples form an indicator's historical distribution).  The results, shown and
explained more precisely in the supplement, are vastly different (worse) than
the original set of results.  In both forecasting and hotspot prediction, the
``fake'' indicator models (with random noise as auxiliary features) offered
essentially no improvement over the pure AR model, which provides a strong
rejection (informally speaking) of the implicit regularization hypothesis.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{fig/hot-1.pdf}
\caption{TODO}
\label{fig:hotspot}
\end{figure}

@@

@@

\subsection{Effects of Leading or Lagging Behavior}

@@

Can axuiliary indicators improve COVID-19 forecasting and hotspot prediction
models?  Our answer, based on analyzing five auxiliary indicators from the  
COVIDcast API (defined using from medical insurance claims, internet-based
surveys, and Google search trends) is undoubtedbly ``yes''. However, there are 
levels of nuance to such an answer that must be explained.  None of the
indicators that we have investigated appear to be the ``silver bullet'' that 
one  might have hoped for, revolutionizing the tractability of the prediction 
problem, rendering it ``easy'' when it was once ``hard'', in the absence of
auxiliary information.  Rather, the gains in accuracy offered by the indicator
models appear to be nontrivial, and consistent across modes of analysis, but
modest---and more about tipping the bias-variance tradeoff in our favor than
providing a consistent leading indication of future events.

[TODO:  Explain how insights from up/down/flat, lagging/leading, etc., are
relevant to this overall arc.  Reiterate the importance of vintage data.  Finish
with something about what's left for future work?  ]

% RJT: some text commented out below.  May be useful for finishing off the
% discussion

% This paper is as much about demonstrating how one might go about answering
% such questions about feature importance in forecasting and hotspot prediction 
% general, as providing an empirically rigorous answer to this precise question
% in the context of the COVID-19 pandemic.
